### 2025-05-14 paper/paper.Rmd
Produce PDF of paper

---
title: "Analyzing Patterns and Causes of TTC Streetcar Delays: January-March 2025"
author: "Allyson Cui"
date: "May 13, 2025"
format: pdf
abstract: "This study examines the Toronto Transit Commission (TTC) streetcar delay data from January to March 2025. By analyzing over 4,000 delay incidents, we identify patterns in delay frequency, location, time, and cause. Our findings reveal that miscellaneous operational issues account for the majority of delays (around 60%), followed by transportation, security, and equipment-related problems. Delays occur consistently throughout the week, with slight variations by day. Time analysis shows peak delay periods in the afternoon (2-3 PM) and evening (8-9 PM). Spatially, Broadview Station experiences the highest number of delays. These insights can inform TTC's resource allocation and preventive maintenance strategies to improve service reliability and passenger experience."
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Code and data available at:** https://github.com/AllysonCui/TTCStreetcarDelay

## Introduction

Public transit systems are essential infrastructure for urban mobility, and their reliability significantly impacts quality of life and economic productivity in cities. In Toronto, the streetcar network is a vital transportation artery serving thousands of commuters daily. However, like any transit system, it experiences service disruptions that affect its reliability [@diab2013].

Delay analysis is crucial for transit authorities to understand where, when, and why disruptions occur. Previous research has demonstrated that identifying patterns in transit delays can lead to more effective resource allocation and preventive maintenance strategies [@barron2013].

For the Toronto Transit Commission (TTC), which operates one of North America's largest streetcar networks, understanding delay patterns can lead to significant improvements in service reliability. According to the TTC's own service standards, maintaining consistent headways (time between vehicles) is a priority for passenger satisfaction [@ttc2024].

This paper analyzes TTC streetcar delay data from January to March 2025, seeking to:
1. Identify spatial patterns in delay occurrences
2. Analyze temporal trends, including time-of-day and day-of-week variations
3. Categorize and quantify the causes of delays
4. Provide data-driven insights for potential service improvements

By systematically examining these aspects, this study aims to contribute to the understanding of transit reliability challenges and inform strategies for enhancing service quality.

## Data

### TTC Streetcar Delay Data

The data used in this analysis comes from the Toronto Open Data Portal's "TTC Streetcar Delay Data" dataset, which provides detailed records of delay incidents across the TTC streetcar network. The dataset is published by the Toronto Transit Commission as part of their open data initiative.

The primary dataset includes the following fields:

- **_id**: Unique identifier for each delay record
- **Date**: The date when the delay occurred (YYYY-MM-DD format)
- **Line**: The streetcar route number and name (e.g., "504 KING")
- **Time**: The time when the delay occurred (24-hour format)
- **Day**: Day of the week
- **Station**: Location description where the delay occurred
- **Code**: TTC delay code (categorized alphanumeric code)
- **Min Delay**: Delay in minutes to the schedule for the following vehicle
- **Min Gap**: Time length in minutes from the vehicle ahead of the following vehicle
- **Bound**: Direction of the streetcar route (E, W, N, S)
- **Vehicle**: Vehicle number

Here's an example record from the dataset:

| _id | Date       | Line        | Time  | Day       | Station             | Code  | Min Delay | Min Gap | Bound | Vehicle |
|-----|------------|-------------|-------|-----------|---------------------|-------|-----------|---------|-------|---------|
| 1   | 2025-01-01 | 504 KING    | 02:10 | Wednesday | KING AND PARLIAMENT | MTSAN | 10        | 20      | W     | 4569    |
| 2   | 2025-01-01 | 506 CARLTON | 02:50 | Wednesday | COLLEGE AND HURON   | MTAFR | 34        | 49      | E     | 4480    |
| 3   | 2025-01-01 | 504 KING    | 03:11 | Wednesday | KING AND QUEEN EAST | MTIE  | 15        | 30      | E     | 4629    |
| 4   | 2025-01-01 | 501 QUEEN   | 05:14 | Wednesday | QUEEN AND DUFFERIN  | MTAFR | 36        | 55      | E     | 4455    |
| 5   | 2025-01-01 | 504 KING    | 06:14 | Wednesday | LESLIE CARHOUSE     | MTVIS | 10        | 20      | E     | 4483    |

The first record indicates a 10-minute delay on the 504 King streetcar line at King and Parliament on Wednesday, January 1, 2025, at 2:10 AM, with a delay code of MTSAN (Unsanitary Vehicle).

### Delay Codes and their Descriptions

A supplementary dataset provides descriptions for the delay codes, with fields:

- **_id**: Unique identifier for each code
- **CODE**: The alphanumeric delay code
- **DESCRIPTION**: Human-readable description of the delay cause

For example:

| _id | CODE  | DESCRIPTION               |
|-----|-------|----------------------------|
| 1   | ETAC  | HVAC                      |
| 2   | ETAR  | ARTICULATION (ALRV)       |
| 3   | ETAX  | AUXILARY POWER SUPPLY     |
| 4   | ETBO  | BODY                      |
| 5   | ETCA  | COMPRESSED AIR            |

The delay codes are categorized with prefixes that indicate the general category of the delay:

- **ET**: Equipment-related delays
- **MT**: Miscellaneous operations-related delays
- **PT**: Plant-related delays (infrastructure)
- **ST**: Security-related delays
- **TT**: Transportation-related delays

It's important to note that the directory on delay codes does not contain explanations for all codes mentioned in the delay data. Specifically, codes that start with 'EF', 'MF', 'SF', 'TF', or 'PF' are not explained in the reference dataset. This limitation means that some delay causes in our analysis remain without detailed descriptions.

### Data Processing and Analysis

The data processing pipeline for this analysis involved several steps:
1. Obtaining the raw data from Toronto's Open Data Portal using their API.
2. Merging the delay incidents with their corresponding code descriptions, handling malformed characters, and standardizing formats.
3. Deriving new features including hour of day and categorical grouping of delay codes.
4. Performing various aggregations to identify patterns across time, space, and delay categories.
5. Creating informative visualizations to represent the findings.

Python was used for all data processing and analysis, with pandas for data manipulation, matplotlib and seaborn for visualization, and specialized libraries like squarify for treemap visualizations. The analysis scripts were structured to follow a consistent pattern of loading, processing, analyzing, and visualizing the data.

## Results

### Spatial Distribution of Delays

The spatial analysis reveals key problem areas in the TTC streetcar network (Figure 1). Three major stations account for 25% of all location-specific delays: Broadview Station (103 delays), Spadina Station (86 delays), and Dundas West Station (75 delays). These are all major transfer points where multiple routes connect.

![Stations with the highest number of delays](../outputs/06-delay_stations.png)

*Figure 1: Top 15 stations with the highest number of delays.*

Terminal loops also show high delay rates, with Sunnyside Loop (48 delays), Humber Loop (46 delays), and Kingston Loop (42 delays) all appearing in the top 10.
The east side of the network appears to have more delay hotspots than the west side, with several major delay locations (Broadview Station, Kingston Loop) in the eastern portion of the network. This may indicate differences in infrastructure quality or operational practices between these areas.

### Temporal Patterns

#### Time of Day Analysis

The hourly delay distribution challenges typical assumptions about when transit is most reliable (Figure 2). Delays peak at 2-3 PM (216 delays) and 8-9 PM (204 delays), not during traditional rush hours.

![Frequency of Delays by Hour of Day](../outputs/05-delay_time_barchart.png)

*Figure 2: Distribution of delay incidents by hour of day.*

The 2-3 PM peak occurs before the afternoon rush, suggesting that preparing for peak service may encounter more problems than handling peak crowds. The 8-9 PM peak happens when supervisor coverage is typically reduced, which may explain the higher delay rates.

Early morning hours (4-6 AM) show consistent delays despite low ridership and traffic. This points to internal issues like vehicle deployment and staff availability at the start of service.

The relatively small variation between hours suggests that delay factors are present throughout the day rather than being strongly tied to specific peak periods.

#### Day of Week Distribution

Delays occur at similar rates across all days of the week (Figure 3), with Saturday showing only slightly more delays (16.1%) than the lowest day, Wednesday (12.8%).

![Delays by Day of Week](../outputs/03-weekday_delays.png)

*Figure 3: Distribution of delays by day of the week.*

This pattern challenges the common practice of reducing resources on weekends. With weekend days accounting for nearly one-third of all delays, and fewer staff typically scheduled on weekends, each delay may have greater impact due to reduced response capacity.

#### Monthly Trends

February shows nearly twice as many delays (1,900) as January (1,000) or March (1,100), as seen in Figure 4.

![Monthly Delay Distribution](../outputs/02-monthly_delays.png)

*Figure 4: Distribution of delays by month in 2025.*

This February spike likely reflects the accumulated effects of winter conditions. While January brings the initial winter weather, February combines sustained cold, built-up snow and ice, and ongoing salt damage to infrastructure. The decrease in March suggests some recovery as conditions improve, though delay levels remain higher than January.

### Delay Causes

#### Categories of Delay Incidents

The analysis of delay types shows that operational issues, not equipment failures, are the main problem (Figure 5).

![Delay Incident Categories](../outputs/04-top_delay_codes.png)

*Figure 5: Treemap showing the distribution of delay incidents by category.*

Miscellaneous Operations [MT] account for 60% of all delays (2,442 incidents), far outweighing equipment problems. Security issues [ST] cause 14% of delays (571 incidents), showing how passenger behavior significantly impacts service.
Equipment failures [ET] (194 incidents, 5%) and infrastructure issues [PT] (67 incidents, 2%) are relatively minor factors, suggesting that vehicle and track reliability are not the main problems.

#### Specific Delay Codes

Looking at specific delay codes reveals more detail about what causes service disruptions (Figure 6).

![Top 15 Delay Incident Codes](../outputs/04-top_delay_codes.png)

*Figure 6: Bar chart showing the top 15 delay incident codes by frequency.*

Just four types of incidents account for over 40% of all delays:

1. **Auto Foul Rail (MTAFR)**: Over 650 incidents where vehicles encounter track obstructions
2. **On Diversion (MTDV**): About 400 incidents where vehicles must detour from normal routes
3. **Disorderly Patron (STDP)**: Over 300 incidents involving passenger behavior issues
4. **Unsanitary Vehicle (MTSAN)**: Over 300 incidents requiring vehicle cleaning

The high rate of Auto Foul Rail incidents points to consistent problems with track-vehicle interfaces. The frequency of Disorderly Patron delays highlights how social factors affect transit reliability.

## Discussion

### Spatial Concentration of Delays

The concentration of delays at major stations such as Broadview, Spadina, and Dundas West suggests that these interchange points may benefit from targeted interventions. These stations serve as connection points between multiple routes and transportation modes, creating complex operational environments. The TTC could consider:

1. Implementing specialized staff training for operators and supervisors at these high-delay locations
2. Reviewing scheduling practices at interchange points to build in appropriate recovery time

The high incidence of delays at loop terminals (Humber, Kingston, Sunnyside) also suggests that turn-around operations may need review. Delays at these points can cascade throughout the system, affecting multiple routes.

### Temporal Insights

The time-of-day analysis reveals several critical periods for delay management:

???

The monthly spike in February aligns with typical winter conditions in Toronto and highlights the need for seasonal preparedness strategies. The TTC might consider:

1. Enhanced winter operational protocols specific to February conditions
2. Preventive maintenance focused on cold-weather vulnerabilities
3. Staffing adjustments to account for the historically higher February delay rates

### Delay Causes and Potential Interventions

The dominance of Miscellaneous Operations (MT) delays, particularly "Auto Foul Rail" (MTAFR) and "On Diversion" (MTDV), points to operational issues as the primary contributor to service disruptions. These operational delays often involve interactions between vehicles and the rail environment, suggesting that:

1. Track maintenance schedules may need review to reduce instances of fouled rails
2. Diversion management procedures could be streamlined to reduce delay impacts

The high incidence of security-related delays (ST category), particularly "Disorderly Patron" (STDP), indicates that passenger behavior significantly impacts service reliability. If more funding is permitted, The TTC might consider:

1. Expanding customer service training for operators to better manage disruptive situations
2. Increasing security presence at high-incident locations identified in the spatial analysis

Equipment-related delays (ET category), while less frequent, still represent significant service disruptions. A preventive maintenance program targeting the most common equipment failures could help reduce these incidents.

### Limitations and Future Research

This analysis has several limitations that should be acknowledged:

1. The three-month time frame (January-March 2025) may not capture seasonal variations throughout the entire year.
2. The lack of explanation for certain delay codes (those beginning with 'EF', 'MF', 'SF', 'TF', or 'PF') limits our understanding of some delay categories.
3. The dataset does not include information on the number of passengers affected by each delay, which would help quantify the service impact.

Future research could address these limitations by:

1. Extending the analysis to a full year of data to capture seasonal patterns
2. Incorporating passenger volume data to weigh delays by their impact
3. Conducting comparative analysis with previous years to identify trends over time
4. Investigating the relationship between weather conditions and specific delay types

## References

### 2025-05-14 scripts/00-retrieve_data.py

@@ -1,10 +1,4 @@
 import requests
-from pathlib import Path
-
-# Find the repository root directory
-current_file = Path(__file__)
-# Go up from current script location to find the root (where the script is in /scripts/)
-repo_root = current_file.parent.parent
 
 # Toronto Open Data is stored in a CKAN instance. It's APIs are documented here:
 # https://docs.ckan.org/en/latest/api/
@@ -30,7 +24,7 @@ for _, resource in enumerate(package["result"]["resources"]):
         resource_dump_data = requests.get(url).text
 
         # Save the CSV data to file
-        save_path = repo_root / "data" / "raw_data" / f"delay_data_{count}.csv"
+        save_path = f"../data/raw_data/delay_data_{count}.csv"
         with open(save_path, 'w') as f:
             f.write(resource_dump_data)
 

### 2025-05-14 scripts/01-clean_data.py

@@ -1,24 +1,14 @@
 import pandas as pd
-from pathlib import Path
-import glob
 
 
 def merge_2025plus_delay_datasets():
 
-    # Find the repository root directory
-    current_file = Path(__file__)
-    # Go up from current script location to find the root
-    repo_root = current_file.parent.parent
-
     # Input file paths
-    first_dataset_path = repo_root / "data" / "raw_data" / "delay_data_0.csv"
-    second_dataset_path = repo_root / "data" / "raw_data" / "delay_data_1.csv"
+    first_dataset_path = "../data/raw_data/delay_data_0.csv"
+    second_dataset_path = "../data/raw_data/delay_data_1.csv"
 
     # Output file path
-    output_path = repo_root / "data" / "analysis_data" / "2025plus_data.csv"
-
-    # Create the output directory if it doesn't exist
-    output_path.parent.mkdir(parents=True, exist_ok=True)
+    output_path = "../data/analysis_data/2025plus_data.csv"
 
     # Read the datasets
     df_delays = pd.read_csv(first_dataset_path)

### 2025-05-14 scripts/02-monthly_delays.py

@@ -1,21 +1,10 @@
 import pandas as pd
 import matplotlib.pyplot as plt
 import numpy as np
-from pathlib import Path
 import calendar
 
-# Set matplotlib backend to a non-interactive one to avoid PyCharm issues
-import matplotlib
-
-matplotlib.use('Agg')  # Use non-interactive backend
-
-# Find the repository root directory
-current_file = Path(__file__)
-# Go up from current script location to find the root
-repo_root = current_file.parent.parent
-
 # Define path to data file
-data_file = repo_root / "data" / "analysis_data" / "2025plus_data.csv"
+data_file = "../data/analysis_data/2025plus_data.csv"
 
 # Load the data
 df = pd.read_csv(data_file)
@@ -163,16 +152,14 @@ ax.set_yticklabels(y_labels)
 # Add gridlines
 ax.grid(axis='y', linestyle='-', alpha=0.2)
 
-# Always add the legend, regardless of the number of years
+# Add the legend, regardless of the number of years
 ax.legend(loc='upper left')
 
 # Adjust layout
 plt.tight_layout()
 
 # Save the figure
-output_dir = repo_root / "outputs"
-output_dir.mkdir(exist_ok=True)
-output_path = output_dir / "02-monthly_delays.png"
+output_path = "../outputs/02-monthly_delays.png"
 plt.savefig(output_path, dpi=300)
 print(f"Figure saved to: {output_path}")
 

### 2025-05-14 scripts/03-weekday_delays.py

@@ -1,13 +1,7 @@
 import pandas as pd
 import matplotlib.pyplot as plt
-from pathlib import Path
-import numpy as np
 
-# Find the repository root directory
-current_file = Path(__file__)
-# Go up from current script location to find the root
-repo_root = current_file.parent.parent
-data_path = repo_root / "data" / "analysis_data" / "2025plus_data.csv"
+data_path = "../data/analysis_data/2025plus_data.csv"
 
 # Load the dataset
 df = pd.read_csv(data_path)
@@ -61,7 +55,7 @@ plt.title('Delays are evenly spread out throughout the week', fontsize=16)
 plt.axis('equal')
 
 # Save the figure to the appropriate location
-save_path = repo_root / "outputs" / "03-weekday_delays.png"
+save_path = output_path = "../outputs/03-weekday_delays.png"
 plt.savefig(save_path, bbox_inches='tight', dpi=300)
 
 # Print only necessary output (matching the style of the reference code)

### 2025-05-14 scripts/04-delay_reason.py

@@ -1,9 +1,7 @@
 import pandas as pd
 import matplotlib.pyplot as plt
-import numpy as np
 import squarify
 import seaborn as sns
-from pathlib import Path
 
 
 def categorize_incidents(df):
@@ -32,7 +30,7 @@ def categorize_incidents(df):
     return df
 
 
-def create_treemap(df, repo_root):
+def create_treemap(df):
     # Count incidents by category
     category_counts = df['Category'].value_counts().to_dict()
 
@@ -64,14 +62,14 @@ def create_treemap(df, repo_root):
     plt.axis('off')
 
     # Save the figure to the appropriate location
-    save_path = repo_root / "outputs" / "04-delay_incidents_treemap.png"
+    save_path = "../outputs/04-delay_incidents_treemap.png"
     plt.savefig(save_path, bbox_inches='tight', dpi=300)
     plt.close()
 
     print("Treemap saved as 'delay_incidents_treemap.png'")
 
 # Function to analyze and create stacked bar chart
-def create_category_bar_chart(df, repo_root):
+def create_category_bar_chart(df):
 
     # Group by code and count occurrences
     code_counts = df['Code'].value_counts().reset_index()
@@ -106,19 +104,14 @@ def create_category_bar_chart(df, repo_root):
     plt.tight_layout()
 
     # Save the figure to the appropriate location
-    save_path = repo_root / "outputs" / "04-top_delay_codes.png"
+    save_path = "../outputs/04-top_delay_codes.png"
     plt.savefig(save_path, bbox_inches='tight', dpi=300)
     plt.close()
 
     print("Bar chart saved as 'top_delay_codes.png'")
 
 
-# Find the repository root directory
-current_file = Path(__file__)
-# Go up from current script location to find the root
-repo_root = current_file.parent.parent
-
-data_path = repo_root / "data" / "analysis_data" / "2025plus_data.csv"
+data_path = "../data/analysis_data/2025plus_data.csv"
 
 # Load the main dataset directly from the specified path
 df = pd.read_csv(data_path)
@@ -127,8 +120,8 @@ df = pd.read_csv(data_path)
 df = categorize_incidents(df)
 
 # Create and save treemap
-create_treemap(df, repo_root)
+create_treemap(df)
 
-create_category_bar_chart(df, repo_root)
+create_category_bar_chart(df)
 
 print("Analysis complete!")

### 2025-05-14 scripts/05-time_delays.py

@@ -2,14 +2,8 @@ import pandas as pd
 import matplotlib.pyplot as plt
 import seaborn as sns
 import numpy as np
-from pathlib import Path
 
-# Find the repository root directory
-current_file = Path(__file__)
-# Go up from current script location to find the root
-repo_root = current_file.parent.parent
-
-data_path = repo_root / "data" / "analysis_data" / "2025plus_data.csv"
+data_path = "../data/analysis_data/2025plus_data.csv"
 
 # Load the data
 data = pd.read_csv(data_path)
@@ -85,7 +79,7 @@ for i, count in enumerate(hour_counts['Count']):
         ax.text(i, count + 1, str(count), ha='center', fontsize=10)
 
 # Save the figure
-save_path = repo_root / "outputs" / "05-delay_time_barchart.png"
+save_path = "../outputs/05-delay_time_barchart.png"
 plt.savefig(save_path, bbox_inches='tight', dpi=300)
 
 print(f"Saved delay times bar chart to file: {save_path}")

### 2025-05-14 scripts/06-location_delays.py

@@ -1,13 +1,7 @@
 import pandas as pd
 import matplotlib.pyplot as plt
-import numpy as np
-from pathlib import Path
 
-# Find the repository root directory
-current_file = Path(__file__)
-# Go up from current script location to find the root
-repo_root = current_file.parent.parent
-data_path = repo_root / "data" / "analysis_data" / "2025plus_data.csv"
+data_path = "../data/analysis_data/2025plus_data.csv"
 
 # Load the data
 data = pd.read_csv(data_path)
@@ -71,7 +65,7 @@ total_records = len(station_counts)
 plt.figtext(0.5, 0.05, f'1 - 15 / {total_records}', ha='center')
 
 # Save the figure
-save_path = repo_root / "outputs" / "06-delay_stations.png"
+save_path = "../outputs/06-delay_stations.png"
 plt.savefig(save_path, bbox_inches='tight', dpi=300)
 
 # Print output

### 2025-05-16 paper/paper.Rmd
Final paper

@@
 
-The primary dataset includes the following fields:
-
-- **_id**: Unique identifier for each delay record
-- **Date**: The date when the delay occurred (YYYY-MM-DD format)
-- **Line**: The streetcar route number and name (e.g., "504 KING")
-- **Time**: The time when the delay occurred (24-hour format)
-- **Day**: Day of the week
-- **Station**: Location description where the delay occurred
-- **Code**: TTC delay code (categorized alphanumeric code)
-- **Min Delay**: Delay in minutes to the schedule for the following vehicle
-- **Min Gap**: Time length in minutes from the vehicle ahead of the following vehicle
-- **Bound**: Direction of the streetcar route (E, W, N, S)
-- **Vehicle**: Vehicle number
-
-Here's an example record from the dataset:
-
-| _id | Date       | Line        | Time  | Day       | Station             | Code  | Min Delay | Min Gap | Bound | Vehicle |
-|-----|------------|-------------|-------|-----------|---------------------|-------|-----------|---------|-------|---------|
-| 1   | 2025-01-01 | 504 KING    | 02:10 | Wednesday | KING AND PARLIAMENT | MTSAN | 10        | 20      | W     | 4569    |
-| 2   | 2025-01-01 | 506 CARLTON | 02:50 | Wednesday | COLLEGE AND HURON   | MTAFR | 34        | 49      | E     | 4480    |
-| 3   | 2025-01-01 | 504 KING    | 03:11 | Wednesday | KING AND QUEEN EAST | MTIE  | 15        | 30      | E     | 4629    |
-| 4   | 2025-01-01 | 501 QUEEN   | 05:14 | Wednesday | QUEEN AND DUFFERIN  | MTAFR | 36        | 55      | E     | 4455    |
-| 5   | 2025-01-01 | 504 KING    | 06:14 | Wednesday | LESLIE CARHOUSE     | MTVIS | 10        | 20      | E     | 4483    |
+| Field Name | Description | Example |
+|------|-------------|------------|
+| _id | Unique identifier for the record | 1 |
+| Date | Date (YYYY/MM/DD) | 2025-01-01 |
+| Line | TTC streetcar line | 504 KING |
+| Time | Time (24h clock) | 02:10 |
+| Day | Name of the day of the week | Wednesday |
+| Station | TTC streetcar station/stop name | KING AND PARLIAMENT |
+| Code | TTC delay code | MTSAN |
+| Min Delay | Delay (in minutes) to the schedule for the following vehicle | 10 |
+| Min Gap | Time length (in minutes) from the vehicle ahead | 20 |
+| Bound | Direction of the route | W |
+| Vehicle | Vehicle number | 4569 |
 
@@
 
-- **_id**: Unique identifier for each code
-- **CODE**: The alphanumeric delay code
-- **DESCRIPTION**: Human-readable description of the delay cause
-
-For example:
-
-| _id | CODE  | DESCRIPTION               |
-|-----|-------|----------------------------|
-| 1   | ETAC  | HVAC                      |
-| 2   | ETAR  | ARTICULATION (ALRV)       |
-| 3   | ETAX  | AUXILARY POWER SUPPLY     |
-| 4   | ETBO  | BODY                      |
-| 5   | ETCA  | COMPRESSED AIR            |
+| Field Name  | Description                         | Example                           |
+|--------------------|----------------------------------|-----------------------------------|
+| _id         | Unique identifier for the code      | 6                                 |
+| CODE        | Delay code                          | ETCE                              |
+| DESCRIPTION | Humanized description for the code  | COMMUNICATION EQUIPMENT<br>(INCLUDES STOP ANNOUNCEMENT) |
 
@@
 The data processing pipeline for this analysis involved several steps:
-1. Obtaining the raw data from Toronto's Open Data Portal using their API.
+
+1. Obtaining the raw data from Toronto's Open Data Portal using their API. Data from previous years are in xlsx format and couldn't be automatically extracted by API.
+
 2. Merging the delay incidents with their corresponding code descriptions, handling malformed characters, and standardizing formats.
+
 3. Deriving new features including hour of day and categorical grouping of delay codes.
+
 4. Performing various aggregations to identify patterns across time, space, and delay categories.
+
 5. Creating informative visualizations to represent the findings.
@@
 The 8-9 PM peak happens when supervisor coverage is typically reduced, which may explain the higher delay rates.
-
-Early morning hours (4-6 AM) show consistent delays despite low ridership and traffic. This points to internal issues like vehicle deployment and staff availability at the start of service.
+Early morning hours (4-6 AM) show consistent delays despite low ridership and traffic.
+This points to internal issues like vehicle deployment and staff availability at the start of service.
 
@@
 
-The high rate of Auto Foul Rail incidents points to consistent problems with track-vehicle interfaces.
-The frequency of Disorderly Patron delays highlights how social factors affect transit reliability.
+This analysis reveals that Toronto's streetcar system struggles most with its operations at the street level.
+Physical obstacles in tracks, necessary route changes due to construction, passenger-related disruptions, and vehicle cleanliness issues collectively create more service problems than mechanical failures.
+This suggests that reliability improvements should focus on street-level operations and passenger management rather than solely pursuing vehicle or infrastructure upgrades.
 
@@
 
-### Spatial Concentration of Delays
+Delays cluster at major stations like Broadview, Spadina, and Dundas West where many routes connect.
+These busy transfer points need special attention, including better staff training and smarter scheduling.
+Loop terminals such as Humber, Kingston, and Sunnyside also show high delay rates, suggesting problems with turn-around operations that can affect the entire system.
 
-The concentration of delays at major stations such as Broadview, Spadina, and Dundas West suggests that these interchange points may benefit from targeted interventions.
-These stations serve as connection points between multiple routes and transportation modes, creating complex operational environments.
-The TTC could consider:
+The times with most delays aren't the usual rush hours.
+Most problems happen from 2-3 PM when service transitions to evening peak, and again from 8-9 PM when fewer supervisors are working.
+Early morning (4-6 AM) also shows many delays despite light traffic, pointing to issues with getting vehicles and crews ready for the day.
+February's higher delay rate shows how winter conditions build up over time, calling for better winter maintenance and possibly more staff during cold months.
 
-1. Implementing specialized staff training for operators and supervisors at these high-delay locations
-2. Reviewing scheduling practices at interchange points to build in appropriate recovery time
+Most delays come from everyday operational issues rather than broken equipment.
+Track blockages and route changes due to construction cause the most disruptions, suggesting we need better track maintenance and smoother handling of diversions.
+Passenger behavior also causes significant delays, showing that better customer management could improve service.
+While equipment problems cause fewer delays, targeted maintenance could still help, especially at trouble spots and during problem times.
 
-The high incidence of delays at loop terminals (Humber, Kingston, Sunnyside) also suggests that turn-around operations may need review.
-Delays at these points can cascade throughout the system, affecting multiple routes.
-
-### Temporal Insights
-
-The time-of-day analysis reveals several critical periods for delay management:
-
-???
-
-The monthly spike in February aligns with typical winter conditions in Toronto and highlights the need for seasonal preparedness strategies. The TTC might consider:
-
-1. Enhanced winter operational protocols specific to February conditions
-2. Preventive maintenance focused on cold-weather vulnerabilities
-3. Staffing adjustments to account for the historically higher February delay rates
-
-### Delay Causes and Potential Interventions
-
-The dominance of Miscellaneous Operations (MT) delays, particularly "Auto Foul Rail" (MTAFR) and "On Diversion" (MTDV), points to operational issues as the primary contributor to service disruptions.
-These operational delays often involve interactions between vehicles and the rail environment, suggesting that:
-
-1. Track maintenance schedules may need review to reduce instances of fouled rails
-2. Diversion management procedures could be streamlined to reduce delay impacts
-
-The high incidence of security-related delays (ST category), particularly "Disorderly Patron" (STDP), indicates that passenger behavior significantly impacts service reliability.
-If more funding is permitted, The TTC might consider:
-
-1. Expanding customer service training for operators to better manage disruptive situations
-2. Increasing security presence at high-incident locations identified in the spatial analysis
-
-Equipment-related delays (ET category), while less frequent, still represent significant service disruptions.
-A preventive maintenance program targeting the most common equipment failures could help reduce these incidents.
-
-### Limitations and Future Research
+## Limitations and Future Research
 
@@
 2. The lack of explanation for certain delay codes (those beginning with 'EF', 'MF', 'SF', 'TF', or 'PF') limits our understanding of some delay categories.
-3. The dataset does not include information on the number of passengers affected by each delay, which would help quantify the service impact.
 
@@
 1. Extending the analysis to a full year of data to capture seasonal patterns
-2. Incorporating passenger volume data to weigh delays by their impact
+2. Incorporating geographical data to better visualize the clusters
 3. Conducting comparative analysis with previous years to identify trends over time

### 2025-05-16 paper/paper.Rmd
Correction on graph
Correction on date

@@
 author: "Allyson Cui"
-date: "May 13, 2025"
+date: "May 16, 2025"
 format: pdf
@@
 
-![Delay Incident Categories](../outputs/04-top_delay_codes.png)
+![Delay Incident Categories](../outputs/04-delay_incidents_treemap.png)
 

### 2025-05-16 scripts/01-clean_data.py
Adjust scripts and README.md.

@@ -1,4 +1,4 @@
-import pandas as pd
+import polars as pl
 
 
 def merge_2025plus_delay_datasets():
@@ -10,28 +10,29 @@ def merge_2025plus_delay_datasets():
     # Output file path
     output_path = "../data/analysis_data/2025plus_data.csv"
 
-    # Read the datasets
-    df_delays = pd.read_csv(first_dataset_path)
-    df_codes = pd.read_csv(second_dataset_path)
+    # Read the datasets using polars
+    df_delays = pl.read_csv(first_dataset_path)
+    df_codes = pl.read_csv(second_dataset_path)
 
     # Rename columns in the codes dataframe for clarity
     df_codes = df_codes.rename(
-        columns={'CODE': 'Code', 'DESCRIPTION': 'Description'})
+        {"CODE": "Code", "DESCRIPTION": "Description"}
+    )
 
     # Clean up the Description column by replacing the malformed characters
-    df_codes['Description'] = df_codes['Description'].str.replace('ГўВЂВ“', '-',
-                                                                  regex=False)
+    df_codes = df_codes.with_columns(
+        pl.col("Description").str.replace_all("ГўВЂВ"", "-")
+    )
 
     # Merge the datasets based on the 'Code' column
-    merged_df = pd.merge(
-        df_delays,
-        df_codes[['Code', 'Description']],
-        on='Code',
-        how='left'
+    merged_df = df_delays.join(
+        df_codes.select(["Code", "Description"]),
+        on="Code",
+        how="left"
     )
 
     # Save the merged dataframe to the output path
-    merged_df.to_csv(output_path, index=False)
+    merged_df.write_csv(output_path)
 
     print(f"Merged dataset saved to {output_path}")
 

### 2025-05-16 scripts/02-monthly_delays.py

@@ -1,4 +1,4 @@
-import pandas as pd
+import polars as pl
 import matplotlib.pyplot as plt
 import numpy as np
 import calendar
@@ -6,30 +6,42 @@ import calendar
 # Define path to data file
 data_file = "../data/analysis_data/2025plus_data.csv"
 
-# Load the data
-df = pd.read_csv(data_file)
+# Load the data with polars
+df = pl.read_csv(data_file)
 
-# Convert the Date column to datetime
-df['Date'] = pd.to_datetime(df['Date'])
-
-# Extract year and month
-df['Year'] = df['Date'].dt.year
-df['Month'] = df['Date'].dt.month
+# Convert the Date column to datetime and extract year and month
+df = df.with_columns([
+    pl.col("Date").str.to_datetime(),
+    pl.col("Date").str.to_datetime().dt.year().alias("Year"),
+    pl.col("Date").str.to_datetime().dt.month().alias("Month")
+])
 
 # Count delays by year and month
-monthly_counts = df.groupby(['Year', 'Month']).size().reset_index(name='Count')
+monthly_counts = (
+    df.group_by(["Year", "Month"])
+    .count()
+    .rename({"count": "Count"})
+)
 
 # Get unique years and months in the data
-years = sorted(monthly_counts['Year'].unique())
-months_present = sorted(monthly_counts['Month'].unique())
+years = sorted(monthly_counts.select("Year").unique().to_series().to_list())
+months_present = sorted(monthly_counts.select("Month").unique().to_series().to_list())
 
 print(f"Years found in data: {years}")
 print(f"Months found in data: {months_present}")
 
 # Find the month with the highest average number of delays
 # Group by month and calculate mean across years
-avg_by_month = monthly_counts.groupby('Month')['Count'].mean()
-max_month_num = avg_by_month.idxmax()
+avg_by_month = (
+    monthly_counts.group_by("Month")
+    .agg(pl.col("Count").mean())
+    .sort("Month")
+)
+
+max_month_row = avg_by_month.filter(
+    pl.col("Count") == pl.col("Count").max()
+).row(0)
+max_month_num = max_month_row[0]
 max_month_name = calendar.month_name[max_month_num]
 
 print(f"Month with highest average delays: {max_month_name} ({max_month_num})")
@@ -44,19 +56,18 @@ fig, ax = plt.subplots(figsize=(12, 6))
[…]

### 2025-05-16 scripts/03-weekday_delays.py

@@ -1,34 +1,54 @@
-import pandas as pd
+import polars as pl
 import matplotlib.pyplot as plt
 
 data_path = "../data/analysis_data/2025plus_data.csv"
 
-# Load the dataset
-df = pd.read_csv(data_path)
+# Load the dataset with polars
+df = pl.read_csv(data_path)
 
 # Group by day and date to get unique days in the dataset
 # First, create a combined Date-Day identifier
-df['Date_Day'] = df['Date'] + '_' + df['Day']
+df = df.with_columns(
+    pl.concat_str([pl.col("Date"), pl.lit("_"), pl.col("Day")]).alias("Date_Day")
+)
 
-# Count total occurrences per day
-day_occurrences = df.groupby('Day')['Date_Day'].nunique()
+# Count unique Date_Day combinations per day to get occurrences
+day_occurrences = (
+    df.group_by("Day")
+    .agg(pl.n_unique("Date_Day"))
+    .rename({"n_unique": "Occurrences"})
+)
 
 # Count delays by day of the week
-day_counts = df['Day'].value_counts()
+day_counts = (
+    df.group_by("Day")
+    .count()
+    .rename({"count": "Count"})
+)
+
+# Join the counts and occurrences
+day_stats = day_counts.join(day_occurrences, on="Day")
 
 # Calculate average delays per day occurrence
-avg_delays_per_day = day_counts / day_occurrences
+day_stats = day_stats.with_columns(
+    (pl.col("Count") / pl.col("Occurrences")).alias("AvgDelaysPerDay")
+)
+
+# Convert to a dictionary for plotting
+day_avg_dict = {row[0]: row[3] for row in day_stats.rows()}
 
 # Calculate percentages for the pie chart
-total_avg = avg_delays_per_day.sum()
-day_percentages = (avg_delays_per_day / total_avg * 100).round(1)
+total_avg = sum(day_avg_dict.values())
+day_percentages = {day: round((value / total_avg * 100), 1) for day, value in day_avg_dict.items()}
 
 # Define colors for each day
 colors = ['#4285F4', '#80CBC4', '#E91E63', '#F57C00', '#FFC107', '#8BC34A', '#673AB7']
 
 # Create a custom order for days of the week
 days_order = ['Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday', 'Saturday', 'Sunday']
-ordered_percentages = day_percentages.reindex(days_order)
+
+# Reorder percentages according to days_order
[…]

### 2025-05-16 scripts/04-delay_reason.py

@@ -1,4 +1,4 @@
-import pandas as pd
+import polars as pl
 import matplotlib.pyplot as plt
 import squarify
 import seaborn as sns
@@ -22,24 +22,42 @@ def categorize_incidents(df):
             return 'Other'
 
     # Create a new column with just the first two characters of the code
-    df['code_prefix'] = df['Code'].str[:2]
+    df = df.with_columns(
+        pl.col("Code").str.slice(0, 2).alias("code_prefix")
+    )
 
     # Apply the mapping function to get categories
-    df['Category'] = df['code_prefix'].apply(map_code_to_category)
+    df = df.with_columns(
+        pl.col("code_prefix").map_dict({
+            'ET': 'Equipment [ET]',
+            'MT': 'Miscellaneous Operations [MT]',
+            'PT': 'Plant [PT]',
+            'ST': 'Security [ST]',
+            'TT': 'Transportation [TT]'
+        }, default='Other').alias("Category")
+    )
 
     return df
 
 
 def create_treemap(df):
     # Count incidents by category
-    category_counts = df['Category'].value_counts().to_dict()
+    category_counts = (
+        df.group_by("Category")
+        .count()
+        .sort("count", descending=True)
+        .collect()
+    )
+
+    # Convert to dictionary for treemap
+    category_dict = {row[0]: row[1] for row in category_counts.rows()}
 
     # Prepare data for treemap
     labels = []
     sizes = []
 
     # Create sorted items for consistent order
-    sorted_items = sorted(category_counts.items(), key=lambda x: x[1],
+    sorted_items = sorted(category_dict.items(), key=lambda x: x[1],
                           reverse=True)
 
     for category, count in sorted_items:
@@ -68,27 +86,36 @@ def create_treemap(df):
 
     print("Treemap saved as 'delay_incidents_treemap.png'")
 
+
 # Function to analyze and create stacked bar chart
 def create_category_bar_chart(df):
-
     # Group by code and count occurrences
-    code_counts = df['Code'].value_counts().reset_index()
-    code_counts.columns = ['Code', 'Count']
+    code_counts = (
+        df.group_by("Code")
+        .count()
+        .sort("count", descending=True)
+        .rename({"count": "Count"})
+    )
 
     # Add description and category
-    code_desc_cat = df[['Code', 'Description', 'Category']].drop_duplicates()
[…]

### 2025-05-16 scripts/05-time_delays.py

@@ -1,50 +1,56 @@
-import pandas as pd
+import polars as pl
 import matplotlib.pyplot as plt
 import seaborn as sns
 import numpy as np
 
 data_path = "../data/analysis_data/2025plus_data.csv"
 
-# Load the data
-data = pd.read_csv(data_path)
+# Load the data with polars
+data = pl.read_csv(data_path)
 
 # Extract hour from the 'Time' column
 # First, ensure the 'Time' column is in string format
-data['Time'] = data['Time'].astype(str)
+data = data.with_columns(
+    pl.col("Time").cast(pl.Utf8).alias("Time")
+)
 
 # Extract hour, handle 'None' values
-def extract_hour(time_str):
-    if time_str == 'None' or pd.isna(time_str):
-        return np.nan
-    time_parts = time_str.split(':')
-    if len(time_parts) >= 2:
-        try:
-            return int(time_parts[0])  # Extract just the hour as an integer
-        except ValueError:
-            return np.nan
-    return np.nan
-
-data['Hour'] = data['Time'].apply(extract_hour)
+data = data.with_columns(
+    pl.when(
+        (pl.col("Time") == "None") | pl.col("Time").is_null()
+    ).then(None).otherwise(
+        pl.col("Time").str.split(":").list.get(0).cast(pl.Int32)
+    ).alias("Hour")
+)
 
 # Create a dataframe with all hours (0-23)
-all_hours = pd.DataFrame({'Hour': range(24)})
+all_hours = pl.DataFrame({"Hour": pl.arange(0, 24, dtype=pl.Int32)})
 
 # Count delays by hour
-hour_counts = data['Hour'].value_counts().reset_index()
-hour_counts.columns = ['Hour', 'Count']
+hour_counts = (
+    data.group_by("Hour")
+    .count()
+    .rename({"count": "Count"})
+)
 
 # Merge with all_hours to ensure we have all 24 hours
-hour_counts = pd.merge(all_hours, hour_counts, on='Hour', how='left').fillna(0)
-hour_counts['Count'] = hour_counts['Count'].astype(int)
+hour_counts = all_hours.join(hour_counts, on="Hour", how="left").fill_null(0)
+
+# Convert "Count" column to integer
+hour_counts = hour_counts.with_columns(
+    pl.col("Count").cast(pl.Int32)
+)
 
 # Create time range strings (e.g., "00:00 - 00:59")
[…]

### 2025-05-16 scripts/06-location_delays.py

@@ -1,26 +1,35 @@
-import pandas as pd
+import polars as pl
 import matplotlib.pyplot as plt
 
 data_path = "../data/analysis_data/2025plus_data.csv"
 
-# Load the data
-data = pd.read_csv(data_path)
+# Load the data with polars
+data = pl.read_csv(data_path)
 
 # Count delays by station
-station_counts = data['Station'].value_counts().reset_index()
-station_counts.columns = ['Station', 'Record Count']
+station_counts = (
+    data.group_by("Station")
+    .count()
+    .rename({"count": "Record Count"})
+    .sort("Record Count", descending=True)
+)
 
 # Handle any None or NaN values
-station_counts = station_counts.fillna({'Station': 'Unknown'})
-
-# Sort by Record Count in descending order
-station_counts = station_counts.sort_values('Record Count', ascending=False)
+station_counts = station_counts.with_columns(
+    pl.when(pl.col("Station").is_null())
+    .then(pl.lit("Unknown"))
+    .otherwise(pl.col("Station"))
+    .alias("Station")
+)
 
 # Display the top 15 stations with most delays
-top_stations = station_counts.head(15)[['Station', 'Record Count']].copy()
+top_stations = station_counts.head(15)
 
-# Make a rank column
-top_stations.insert(0, '#', range(1, len(top_stations) + 1))
+# Create a rank column
+top_stations = top_stations.with_row_index(name="#", offset=1)
+
+# Convert to pandas for matplotlib table
+top_stations_pd = top_stations.to_pandas()
 
 # Set up the figure and plot
 plt.figure(figsize=(12, 8))
@@ -32,12 +41,11 @@ ax.yaxis.set_visible(False)
 
 # Create the table
 table = plt.table(
-    cellText=top_stations.values,
+    cellText=top_stations_pd.values,
     colLabels=['#', 'Station', 'Record Count'],
     cellLoc='center',
     loc='center',
     colWidths=[0.1, 0.6, 0.3]
-    # Adjusted widths for the wider station column
 )
 
 # Style the table
@@ -57,8 +65,7 @@ for i, key in enumerate(table._cells):
         cell.set_facecolor('#f0f0f0')
 
 # Add title
-plt.title('Stations with the highest number of delays:', fontsize=14,
-          pad=20)
+plt.title('Stations with the highest number of delays:', fontsize=14, pad=20)
 
 # Add pagination indicator at the bottom
 total_records = len(station_counts)
[…]

### 2025-05-16 paper/paper.Rmd
Updates on uv and polars

@@
 date: "May 16, 2025"
-format: pdf
+format: 
+  pdf:
+    keep-tex: false
+    keep-md: false
+    clean: true
 abstract: "This study examines the Toronto Transit Commission (TTC) streetcar delay data from January to March 2025.

### 2025-05-16 scripts/01-clean_data.py

@@ -21,7 +21,7 @@ def merge_2025plus_delay_datasets():
 
     # Clean up the Description column by replacing the malformed characters
     df_codes = df_codes.with_columns(
-        pl.col("Description").str.replace_all("ГўВЂВ"", "-")
+        pl.col("Description").str.replace_all('ГўВЂВ"', "-")
     )
 
     # Merge the datasets based on the 'Code' column

### 2025-05-16 scripts/02-monthly_delays.py

@@ -19,7 +19,7 @@ df = df.with_columns([
 # Count delays by year and month
 monthly_counts = (
     df.group_by(["Year", "Month"])
-    .count()
+    .len()
     .rename({"count": "Count"})
 )
 

### 2025-05-16 scripts/03-weekday_delays.py

@@ -15,14 +15,13 @@ df = df.with_columns(
 # Count unique Date_Day combinations per day to get occurrences
 day_occurrences = (
     df.group_by("Day")
-    .agg(pl.n_unique("Date_Day"))
-    .rename({"n_unique": "Occurrences"})
+    .agg(pl.col("Date_Day").n_unique().alias("Occurrences"))  # Directly name the column
 )
 
 # Count delays by day of the week
 day_counts = (
     df.group_by("Day")
-    .count()
+    .len()
     .rename({"count": "Count"})
 )
 

### 2025-05-16 scripts/04-delay_reason.py

@@ -28,7 +28,7 @@ def categorize_incidents(df):
 
     # Apply the mapping function to get categories
     df = df.with_columns(
-        pl.col("code_prefix").map_dict({
+        pl.col("code_prefix").replace({
             'ET': 'Equipment [ET]',
             'MT': 'Miscellaneous Operations [MT]',
             'PT': 'Plant [PT]',
@@ -44,9 +44,8 @@ def create_treemap(df):
     # Count incidents by category
     category_counts = (
         df.group_by("Category")
-        .count()
-        .sort("count", descending=True)
-        .collect()
+        .len()
+        .sort("len", descending=True)
     )
 
     # Convert to dictionary for treemap
@@ -89,12 +88,17 @@ def create_treemap(df):
 
 # Function to analyze and create stacked bar chart
 def create_category_bar_chart(df):
+    # Fill any null values in the Description column with a placeholder
+    df = df.with_columns(
+        pl.col("Description").fill_null("Unknown")
+    )
+
     # Group by code and count occurrences
     code_counts = (
         df.group_by("Code")
-        .count()
-        .sort("count", descending=True)
-        .rename({"count": "Count"})
+        .len()
+        .sort("len", descending=True)
+        .rename({"len": "Count"})
     )
 
     # Add description and category
@@ -116,7 +120,11 @@ def create_category_bar_chart(df):
 
     # Add descriptions as annotations
     for i, (_, row) in enumerate(top_codes_pd.iterrows()):
-        plt.annotate(f"{row['Description'][:]}",
+        # Handle any None values in Description
+        description = row['Description'] if row[
+                                                'Description'] is not None else "Unknown"
+
+        plt.annotate(f"{description}",
                      xy=(i, row['Count']),
                      xytext=(0, 5),
                      textcoords="offset points",

### 2025-05-16 scripts/05-time_delays.py

@@ -23,27 +23,30 @@ data = data.with_columns(
     ).alias("Hour")
 )
 
-# Create a dataframe with all hours (0-23)
-all_hours = pl.DataFrame({"Hour": pl.arange(0, 24, dtype=pl.Int32)})
+# Create a dataframe with all hours (0-23) using list and explicitly set dtype to Int32
+all_hours = pl.DataFrame({"Hour": pl.Series(range(24), dtype=pl.Int32)})
 
 # Count delays by hour
 hour_counts = (
     data.group_by("Hour")
-    .count()
-    .rename({"count": "Count"})
+    .len()  # Use len() instead of count() which is deprecated
+    .rename({"len": "Count"})
 )
 
 # Merge with all_hours to ensure we have all 24 hours
-hour_counts = all_hours.join(hour_counts, on="Hour", how="left").fill_null(0)
+# Explicitly set coalesce=True to address the deprecation warning
+hour_counts = all_hours.join(hour_counts, on="Hour", how="left", coalesce=True).fill_null(0)
 
 # Convert "Count" column to integer
 hour_counts = hour_counts.with_columns(
     pl.col("Count").cast(pl.Int32)
 )
 
-# Create time range strings (e.g., "00:00 - 00:59")
+# Create time range strings (e.g., "00:00 - 00:59") using string formatting
+# Simpler approach using string concatenation instead of pl.format
 hour_counts = hour_counts.with_columns(
-    pl.format("{:02d}:00 - {:02d}:59", pl.col("Hour"), pl.col("Hour")).alias("Time Range")
+    (pl.col("Hour").cast(pl.Utf8).str.zfill(2) + ":00 - " +
+     pl.col("Hour").cast(pl.Utf8).str.zfill(2) + ":59").alias("Time Range")
 )
 
 # Sort by Hour

### 2025-05-16 scripts/06-location_delays.py

@@ -9,8 +9,8 @@ data = pl.read_csv(data_path)
 # Count delays by station
 station_counts = (
     data.group_by("Station")
-    .count()
-    .rename({"count": "Record Count"})
+    .len()  # Use len() instead of count() which is deprecated
+    .rename({"len": "Record Count"})  # Updated column name
     .sort("Record Count", descending=True)
 )
 
